Graph-Based Neural Reconstruction from Morphological Skeletons

### 2017-11-15 papers/cvpr2018/0_introduction.tex
paper: BM updates

%\donglai{usually, top-down means split.. but we are also bottom-up, but with different representation}
The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{kasthuri2015saturated,haehn2017scalable}. Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) have allowed researchers to produce terabytes of image data every hour~\cite{hildebrand2017whole}. It is not feasible for domain experts to manually reconstruct thousands or millions of neurons from this vast amount of image data~\cite{haehn2014design}. State-of-the-art automatic reconstruction approaches use pixel-based segmentation with convolutional neural networks (CNNs) followed by agglomeration strategies~\cite{seymour2016rhoananet,nunez2014graph,parag2017anisotropic,zlateski2015image,lee2015recursive,ronneberger2015u}. %TODO DOUBLE CHECK REFERENCES
These \textit{bottom-up pixel-based} algorithms produce excellent results but still fall short of acceptable error rates for large volumes.

@@

Our approach operates at a level of abstraction above existing pixel-based methods. This allows us to leverage both local and global information to produce more accurate reconstructions. Because it uses agglomerated regions our classifier is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining. Using the 3D graph of the segmentation allows us to enforce global constraints on the reconstruction that closely follow the underlying biology, such as limiting the angles of neuron branchings to the biologically plausible range. We can also use local information such as shape priors to weight the edges in our graph. This dual approach of assessing local decisions in a global context yields accuracy improvements over the existing reconstruction methods.

This work makes the following following contributions: (1) a novel top-down graph-based method for neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors under global constraints; (3) an empirical evaluation of our method on several connectomics datasets; (4) our method yields improved performance over the state-of-the-art on \FIX{X out of Y} datasets, and competitive results on the remaining data. On average, we improve the Variation of Information (VI) metric across all datasets by \FIX{X} percent without drastically increasing the running time.

>>>>

%\donglai{usually, top-down means split.. but we are also bottom-up, but with different representation}
The field of connectomics is concerned with reconstructing the wiring diagram of the brain at nanometer resolutions to enable new insights into the workings of the brain~\cite{haehn2017scalable,kasthuri2015saturated}. Recent advancements in image acquisition using multi-beam serial-section electron microscopy (sSEM) have allowed researchers to produce terabytes of image data every hour~\cite{hildebrand2017whole}. It is not feasible for domain experts to manually reconstruct thousands or millions of neurons from this vast amount of image data~\cite{haehn2014design}. State-of-the-art automatic reconstruction approaches use pixel-based segmentation with convolutional neural networks (CNNs) followed by agglomeration strategies~\cite{seymour2016rhoananet,lee2015recursive,nunez2014graph,parag2017anisotropic,ronneberger2015u,zlateski2015image}. %TODO DOUBLE CHECK REFERENCES
These \textit{bottom-up pixel-based} algorithms produce excellent results but still fall short of acceptable error rates for large volumes.

@@

Our approach operates at a level of abstraction above existing pixel-based methods. This allows us to leverage both local and global information to produce more accurate reconstructions. Since it uses agglomerated regions, our classifier is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic image data without retraining. Using the 3D graph of the segmentation allows us to enforce global constraints on the reconstruction that closely follow the underlying biology. We can also use local information such as shape priors to weight the edges in our graph. This dual approach of assessing local decisions in a global context yields accuracy improvements over the existing reconstruction methods.

This work makes the following following contributions: (1) a novel top-down graph-based method for neural reconstruction of connectomics data; (2) a region-based CNN classifier to detect errors under global constraints; (3) an empirical evaluation of our method on several connectomics datasets; (4) our method yields improved performance over the state-of-the-art on three testing datasets. On average, we improve the Variation of Information (VI) metric across all datasets by \FIX{X} percent without drastically increasing the running time.

### 2017-11-15 papers/cvpr2018/2_method.tex
paper: BM updates

\donglai{need some overview sentence}
%We generate nodes $N$ and edges $E$ to apply a graph-based optimization strategy for segmentation.
%In addition, these edges receive non-negative weights.

@@

First, we extract a skeleton from each segment using the TEASER algorithm~\cite{sato2000teasar,zhao2014automatic}. Fig.~\ref{fig:skeletonization} shows an example of two extracted skeletons (in white) of two segments from the label volume. These skeletons consist of a sequence of \textit{joints}, i.e., locations that are a local maximum distance from the segment boundary, with line segments connecting successive joints. We prune the joints that are within $t_{jnt} = 50$ voxels of each other to reduce unnecessary branching. For the purposes of our algorithm, joints that have only one connected neighbor are referred to as \textit{endpoints}. Many of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). We will make use of this fact to merge segments with the following two-pass heuristic.

@@

We extract a cubic region of interest around each endpoints $e$ in $\mathbb{S}_c$ as input to the CNN. These regions of interest provide the local information for the neural network to predict which neighboring segments belong to the same neuron. The CNN receives three input channels for every voxel in the region of interest around segments $l_1$ and $l_2$. The input in all of the channels is in the range $\{-0.5, 0.5\}$. The first channel is $0.5$ only if the corresponding voxel has label $l_1$. The second channel is $0.5$ only if the corresponding voxel has label $l_2$. The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.

@@

\begin{figure*}[t]
	\centering
	\includegraphics[width=0.95\linewidth]{figures/architecture.png}
	\caption{The architecture for our classification CNN uses double convolutions followed by max pooling. The number of filters doubles each layer, with a final fully connected layer and sigmoid activation function.}
	\label{fig:architecture}
\end{figure*}

>>>>

%We generate nodes $N$ and edges $E$ to apply a graph-based optimization strategy for segmentation.
%In addition, these edges receive non-negative weights.

@@

First, we extract a skeleton from each segment using the TEASER algorithm~\cite{sato2000teasar,zhao2014automatic}. Fig.~\ref{fig:skeletonization} shows an example of two extracted skeletons (in black) of two segments from the label volume. These skeletons consist of a sequence of \textit{joints}, i.e., locations that are a local maximum distance from the segment boundary, with line segments connecting successive joints. We prune the joints that are within $t_{jnt} = 50$ voxels of each other to reduce unnecessary branching. For the purposes of our algorithm, joints that have only one connected neighbor are referred to as \textit{endpoints}. Many of the segments that are erroneously split have nearby endpoints  (Fig.~\ref{fig:merge_candidates}). We make use of this fact to merge segments with the following two-pass heuristic.

@@

We extract a cubic region of interest around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. These regions of interest provide the local information for the neural network to predict which neighboring segments belong to the same neuron. The CNN receives three input channels for every voxel in the region of interest around segments $l_1$ and $l_2$. The input in all of the channels is in the range $\{-0.5, 0.5\}$. The first channel is $0.5$ only if the corresponding voxel has label $l_1$. The second channel is $0.5$ only if the corresponding voxel has label $l_2$. The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.

@@

\begin{figure*}[t]
	\centering
	\includegraphics[width=0.85\linewidth]{figures/architecture.png}
	\caption{The architecture for our classification CNN uses double convolutions followed by max pooling. The number of filters doubles each layer, with a final fully connected layer and sigmoid activation function.}
	\label{fig:architecture}
\end{figure*}

### 2017-11-15 papers/cvpr2018/4_results.tex
paper: BM updates

@@

We evaluate the performance of the different methods using the split version of variance of information~\cite{meila2003comparing} (VI-s). Given a ground truth labeling $GT$ and our automatically reconstructed segmentation $SG$.
Over and under-segmentation are quantified by the conditional entropy $H(GT | SG)$ and $H(SG | GT)$, respectively. Since we are measuring the entropy between two clusterings, better VI scores are closer to the origin.

In Fig.~\ref{fig:variation-of-information}, we show VI scores for the input over-segmentation of the Kasthuri data \hp{check} at different thresholds of agglomeration (green) compared to our method (red). We also show the comparison to an oracle (blue) that correctly partitions the graph from our algorithm based on ground truth. \hp{add discussion of the results}

%The oracles sees the graph from our algorithm and correctly partitions the graph based on the ground truth.
%TODO{Our methods decreases the split variation of information by a factor of XX\% and only increases the merge variation of information by a factor of XX\% on the test datasets.}

Fig.~\ref{fig:positive-results} shows successful merges on the Kasthuri Vol. 2. Several of these successful examples combine multiple consecutive segments. The third example in the figure shows the correction of an over-segmented dendrite.
Fig.~\ref{fig:negative-results} shows some failure cases. In two of these examples the algorithm correctly predicted several merges and made one error.
In the third example a catastrophic merge error \hp{check} in the initial segmentation propagated to the output. We now analyze how each major component of our method contributes to the final result.

@@

Equally important is the number of split errors that remain after pruning. These are the locations that we want to merge to create a more accurate reconstruction. For every dataset, the number of positive candidates \hp{what is a positive candidate?} remains relatively equal. \hp{equal to what?}
However, since our heuristic does not enforce an adjacency constraint of two regions when constructing edges in the graph, the difference does not indicate the number of examples excluded by pruning. In fact, our method finds a number of examples that are non-adjacent. Figure \ref{fig:skeleton-results} shows two example segments with split errors, one that our algorithm missed (top) and one that it identified (bottom), despite the fact that the split segments are not adjacent.

@@

Table \ref{table:classification} shows the precision and recall rates for all of the datasets. We can train the CNN on anisotropic data and get good results on isotropic data. \hp{add more discussion}

\begin{table}[h]
	\centering
	\begin{tabular}{c c c c} \hline
		\textbf{Dataset} & \textbf{Precision} & \textbf{Recall} & \textbf{Accuracy} \\ \hline
		Kasthuri Training & 0.919 & 0.936 & 0.974 \\
		Kasthuri Testing & 0.737 & 0.717 & 0.907 \\
		FlyEM Vol. 1 & 0.796 & 0.478 & 0.862 \\
		FlyEM Vol. 2 & 0.762 & 0.422 & 0.810 \\ \hline
	\end{tabular}
	\caption{Precision and recall rates for the training and three test datasets.}
	\label{table:classification}
\end{table}

@@

Applying a graph-based optimization strategy increases the accuracy over the CNN alone. In our test scores, we note an increase in precision on all datasets. For our segmentation problems we prefer a higher precision since it is more difficult to correct merge-errors. Table \ref{table:multicut} shows the changes in precision, recall, and accuracy for all three datasets in relation to the CNN without graph-based optimization. \hp{add more discussion}

>>>>

We evaluate the performance of the different methods using the split version of variance of information (VI)~\cite{meila2003comparing}. 
Given a ground truth labeling $GT$ and our automatically reconstructed segmentation $SG$, over and undersegmentation are quantified by the conditional entropy $H(GT | SG)$ and $H(SG | GT)$, respectively. Since we are measuring the entropy between two clusterings, better VI scores are closer to the origin.

@@

In Fig.~\ref{fig:variation-of-information}, we show the VI results of NeuroProof on the Kasthuri and FlyEM data at varying thresholds of agglomeration (green). 
The green circle represents the variation of information for our input segmentation (a threshold of 0.3 for all datasets). 
Our results are in red.  
We show the comparison to an oracle (blue) that correctly partitions the graph from our algorithm based on ground truth.

Our algorithm improves the accuracy of the input segmentation on every dataset, reducing the VI split score on average by \FIX{X\%} and only increasing the VI merge score by \FIX{X\%}. 
Scores closer to the origin are better for this metric, and in every instance we are below the green curve. 
We see significant improvements on the Kasthuri datasets (VI split reduction of \FIX{X\%} and \FIX{X\%} on the training and testing datasets respectively) and slightly more modest improvements on the FlyEM datasets (reduction of \FIX{X\%} and \FIX{X\%}).
However, our baseline, NeuroProof performs much better on the FlyEM datasets reducing the potential improvement. 
Isotropic datasets are easier to segment using state-of-the-art region-based methods than anisotropic ones~\cite{plaza2014annotating}.
Thus there is less room for improvement on these datasets.

Fig.~\ref{fig:positive-results} shows successful merges on the Kasthuri Vol. 2 dataset. 
Several of these examples combine multiple consecutive segments that span the volume. 
In the third example we correct the oversegmentation of a dendrite.
Fig.~\ref{fig:negative-results} shows some failure cases. 
In two of these examples the algorithm correctly predicted several merges but made one error.
In the third example a merge error in the initial segmentation propagated to our output. 
We now analyze how each major component of our method contributes to this final result.

@@

Equally important is the number of split errors that remain after pruning. These are the locations that we want to merge to create a more accurate reconstruction. For every dataset, the number of true split errors remains constant before and after pruning.
However, since our heuristic does not enforce an adjacency constraint of two regions when constructing edges in the graph, the difference does not indicate the number of examples excluded by pruning. In fact, our method finds a number of examples that are non-adjacent. Figure \ref{fig:skeleton-results} shows two example segments with split errors, one that our algorithm missed (top) and one that it identified (bottom), despite the fact that the split segments are not adjacent.
Of the successful examples in Figure \ref{fig:positive-results}, the second and fourth groupings contain pairs of non-adjacent segments that merge with our algorithm.

@@

Figure \ref{fig:receiver-operating-characteristic} shows the receiver operating characteristic (ROC) curve for all of the datasets. 
We train our CNN using one of the Kasthuri volumes and test using the other three datasets.
Since our CNN only takes as input a region of the label volume we can train on an anisotropic dataset and infer on an isotropic one. 
This provides a major benefit given the time-intensive task of manually generating ground truth data at various resolutions.

As evidenced by the ROC curve, the test results on the Kasthuri data are better than the results for FlyEM.
We believe this is in part because of the differences in the datasets (i.e., isotropy and $xy$ resolution). 
To test this hypothesis, we also evaluate the performance of the FlyEM datasets when the network trains on FlyEM Vol. 1 and infers on FlyEM Vol. 2.\footnote{Since the FlyEM datasets have significantly fewer examples, we initialize the network with the weights from the Kasthuri training and have an initial learning rate of $10^{-4}$.}
The dotted lines in the figure represent these tests. 
There is a slight performance increase for FlyEM Vol. 2.
However, the FlyEM datasets have reasonable results when the network is trained by the Kasthuri data, and the results outside of this section follow from that setup.

\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{./figures/receiver-operating-characteristic.jpg}
	\caption{The receiver operating characteristic for all four datasets.}
	\label{fig:receiver-operating-characteristic}
\end{figure}

@@

The graph-based optimization strategy increases our correction accuracy over using just the CNN. 
In particular, the precision increases on each dataset, although the recall decreases on all but one of the datasets. 
Since it is more difficult to correct merge errors than split errors, it is often desirable to sacrifice recall for precision. 
Table \ref{table:multicut} shows the changes in precision, recall, and accuracy for all four datasets compared to the CNN. 
Over the three testing datasets, applying a graph-based partitioning strategy reduced the number of merge errors by \FIX{X}, \FIX{Y}, and \FIX{Z}.

### 2017-11-15 papers/cvpr2018/5_conclusions.tex
paper: BM updates

We present a novel method for reconstructing neuronal processes in connectomics datasets. We extend existing pixel-based reconstruction strategies using a morphological 3D skeleton graph and show significant accuracy improvement on two datasets from two different species. Our approach enables enforcing domain-constraints on the resulting segmentation. We believe this approach extends beyond connectomics to other domains with different constraints.
\hp{add more future work}

>>>>

We present a novel method for reconstructing neuronal processes in connectomics datasets. 
We extend existing region-based reconstruction strategies using a morphological 3D skeleton graph and show significant accuracy improvement on two datasets from two different species. 
Our approach enables enforcing domain-constraints on the resulting segmentation. 
We believe this approach extends beyond connectomics to other domains with different constraints.

There is significant room for additional research related to this topic.
We can still extract a lot more information from the skeletons of each segment.
Currently these skeletons identify potential merge locations.
However, these representations can be used to identify improperly merged segments.
Additionally, we can further improve our results by tweaking the partitioning cost functions to more closely match the underlying biology.
One such extension might take the outputs of a synaptic detector to enforce the constraint that a given segment only has post- or pre-synaptic connections.

### 2017-11-15 papers/cvpr2018/paper.tex
paper: BM updates

%%%%%%%%% ABSTRACT
\begin{abstract}
Advancements in electron microscopy image acquisition have created massive connectomics datasets in the terabyte range that make manual reconstruction of neuronal structures infeasible. 
Current state-of-the-art automatic methods segment neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions. 
However, these approaches widely neglect global geometric properties that are inherent in the graph structure of neural wiring diagrams.
In this work, we follow bottom-up pixel-based reconstruction by a top-down graph-based method to more accurately approximate neural pathways. 
Using the membrane labels of the pixel-based segmentation we first generate skeletons in 3D.
Using this 3D graph we train automatic classifiers for shape description to detect impossible neural pathways by looking at geometric properties.
We then apply efficient graph-based optimization strategies to improve the segmentation labels. We demonstrate the performance of our approach on multiple real-world connectomics datasets with average variation of information improvement of \FIX{$X\times$}. 
%The paper provides insights into learning shape features for top-down graph-based neuron reconstruction.
\end{abstract}
%%%%%%%%	% BODY TEXT

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
Advancements in electron microscopy image acquisition have created massive connectomics datasets in the terabyte range that make manual reconstruction of neuronal structures infeasible. 
Current state-of-the-art automatic methods segment neural membranes at the pixel level followed by agglomeration methods to create full neuron reconstructions. 
However, these approaches widely neglect global geometric properties that are inherent in the graph structure of neural wiring diagrams.
In this work, we follow bottom-up pixel-based reconstruction by a top-down graph-based method to more accurately approximate neural pathways. 
Using the membrane labels of the pixel-based segmentation we first generate skeletons in 3D.
Using this 3D graph we train automatic classifiers for shape description to detect impossible neural pathways by looking at geometric properties.
We then apply efficient graph-based optimization strategies to improve the segmentation labels. We demonstrate the performance of our approach on multiple real-world connectomics datasets with an average variation of information improvement of \FIX{$X\times$}. 
%The paper provides insights into learning shape features for top-down graph-based neuron reconstruction.
\end{abstract}
%%%%%%%%	% BODY TEXT
